Interpretable Machine Learning

### 2019-01-14 chunks/story-wines.Rmd
continues slides

##  {.emphasizedabit}
A startup wants you to predict wine quality from its chemical composition.

# Let's predict wine quality

Free online dataset about red and white variants of the Portuguese "Vinho Verde" wine.

Target: Quality from 1 to 10
Features: Acidity, alcohol ...

@@

- fixed acidity
- volatile acidity
- citric acid
- residual sugar
- chlorides
- free sulfur dioxide
- total sulfur dioxide
- density
- pH
- sulphates
- alcohol
- quality based on sensory data (0-10)

@@

10x CV
Test linear regression model, decision tree, random forest

@@

preds = pred$predict(wine)
preds$actual = wine$quality
ggplot(preds, aes(x = actual, y = .prediction, group = actual)) + 
  geom_violin(aes(fill = ..n..)) +
  scale_x_continuous("Actual quality (jittered)", 
    labels = 1:10, breaks = 1:10) + 
  scale_y_continuous("Predicted quality", labels = 1:10, breaks = 1:10) + 
  scale_fill_gradient(low = "white", high = "darkblue", guide = "none")
```

@@

@@

@@

## How do features affect predictions?
Method: Partial dependence plot, individual conditional expectation curves
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'pdp+ice')
plot(eff)
```

## Interactions with type of wine?
Method: H-statistic for interactions

```{r}
inter = Interaction$new(pred)
plot(inter)
```

## Interactions of alcohol?
Method: H-statistic for interactions

```{r}
inter.a = Interaction$new(pred, "alcohol")
plot(inter.a)
```

## Rule of thumb for wine quality?
Method: Surrogate Model

```{r surrogate}
tree = TreeSurrogate$new(pred, maxdepth  = 2)
plot(tree$tree)
```

@@

A customer gets a really bad prediction. What was the reason?

@@

min_pred = min(predictions)
worst = wine[which(predictions == min_pred),]

# explain with shapley
knitr::kable(t(worst))
# explain with cf?

min_pred

```

@@

## The best wine

```{r best-wine}
# find bad wine in data
max_pred = max(predictions)
best = wine[which(predictions == max_pred),]

# explain with shapley
knitr::kable(t(best))
# explain with cf?

max_pred

```

## The best wine
Technique: Shapley Values

```{r}
shap = Shapley$new(pred, x.interest = best)
plot(shap)
```

>>>>

##  {.emphasizedabit}
You want to predict wine quality from its physicochemical properties.

## {.center data-background=../images/wine-features.jpg data-background-size=contain}

## Step 1: Find data
Free dataset of red and white "Vinho Verde" wine.

```{r}
knitr::include_graphics("../images/minho.png")
```

Three sensory assessors (using blind tastes), which graded the wine in a scale that ranges from
0 (very bad) to 10 (excellent)

@@

@@

Compare different models with 10x CV
- Linear regression model
- Decision tree
- Random forest

@@

preds = pred$predict(wine)
preds$actual = wine$quality
ggplot(preds, aes(x = actual, y = .prediction, group = actual)) + 
  geom_violin(aes(fill = ..n..)) +
  scale_x_continuous("Actual quality", 
    labels = 1:10, breaks = 1:10) + 
  scale_y_continuous("Predicted quality", labels = 1:10, breaks = 1:10) + 
  scale_fill_gradient(low = "white", high = "darkblue", guide = "none")
```

@@

TODO: Slide explaining permutation feature importance

@@

TODO: Slide explaining ALE plots

## How do features affect predictions?
Method: Accumulated Local Effects
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'ale')
plot(eff)
```

## How do features affect predictions?
Method: Accumulated Local Effects
```{r}
eff = FeatureEffect$new(pred, "volatile.acidity", method = 'ale')
plot(eff)
```

@@

## Interactions between alcohol and volatile acidity?

```{r}
inter = FeatureEffect$new(pred, feature = c("alcohol", "volatile.acidity"))
plot(inter)
```

## Rule of thumb for wine quality?

```{r surrogate}
library(partykit)
tree = TreeSurrogate$new(pred, maxdepth  = 2)
node_inner2 = node_inner(tree$tree, pval = FALSE, id=FALSE)
node_terminal2 = node_boxplot(tree$tree, id = FALSE)
plot(tree$tree, inner_panel = node_inner2, terminal_panel = node_terminal2)

rsquared = tree$r.squared
```

The tree explains `r sprintf("%.2f%s", 100 *  rsquared, "%")` of the black box prediction variance.

@@

@@

min_pred = min(predictions)
worst = wine[which(predictions == min_pred),]
```

TODO: Image of really bad wine

Predicted quality: `r predictions[which(predictions == min_pred),]`

TODO: Slide to explain Shapley value

@@

## What needs to change?

TODO: Slide to explain counterfactuals

## Counterfactual explanation

How do we get the wine above predicted quality of 5?

```{r}
worst

worst2 = function(){
  w = worst
  w$volatile.acidity = 0.2
  w
}

pred$predict(worst2())[[1]]

worst3 = function(){
  w = worst
  w$alcohol = 13
  w$volatile.acidity = 1
  w
}

pred$predict(worst3())[[1]]
```

TODO: Image

- Decreasing volatile acidity to 0.2 yields predicted quality of `r sprintf('%.2f', pred$predict(worst2())[[1]])`
- Decreasing volatile acidity to 1.0 and increasing alcohol to 13% yields predicted quality of `r sprintf('%.2f', pred$predict(worst3())[[1]])`

### 2019-01-14 chunks/tools.Rmd
continues slides

@@

TODO: PDP gif

@@

TODO: Feature importance figure

@@

## Example-based Methods

## Example-focused Methods
```{r, echo=FALSE}
knitr::include_graphics("../images/data-box.png")
```

<div class="notes">
- Prototypes: What is the prototype of a certain class
- Criticisms: Which data points are unusual for a certain class. 
- Counterfactuals: What do I need to change for an instance to change the prediction?
- Anchors: What features do I have to anchor to keep prediction the same?
</div>

>>>>

##

TODO: Overview (drawing)

- global vs. local
- interpretable model vs. post-hoc
- model-specific vs. model-agnostic
- data type image vs. text vs. tabular

@@

TODO: Drawing of feature effect

@@

TODO: Drawing of importance

@@
